---
title: "Reproducible Research: Peer Assessment 1"
author: "Brice Hoareau"
date: "24 January 2016"
output: 
  html_document: 
    keep_md: yes
---

## Loading and preprocessing the data

Load the data from "activity.csv" and inspect:
```{r}
activitydata <- read.csv("activity.csv")
summary(activitydata)
str(activitydata)
```

No preprocessing necessary.

## What is mean total number of steps taken per day?

Calculate the total number of steps taken per day:
```{r results='asis'}
library(xtable)
dailysteps <- aggregate(formula = steps ~ date, data = activitydata, FUN = sum, na.action = na.omit)
print(xtable(dailysteps), type = "html")
```

```{r}
hist(dailysteps$steps, breaks = 8)
```

Histogram of the total number of steps taken each day.

```{r}
dailymean <- mean(dailysteps$steps)
dailymedian <- median(dailysteps$steps)
```

The mean and median of the total number of steps taken per day are __`r dailymean`__ and __`r dailymedian`__ respectively.

## What is the average daily activity pattern?

Time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days:
```{r}
avgintervalsteps <- aggregate(formula = steps ~ interval, data = activitydata, FUN = mean, na.action = na.omit)
plot(avgintervalsteps$interval, avgintervalsteps$steps, type = "l")
```

The 5-minute interval, on average across all the days in the dataset, that contains the maximum number of steps is:
```{r}
avgintervalsteps[row.names(avgintervalsteps) == which.max(avgintervalsteps$steps),]$interval
```

## Imputing missing values

The total number of missing values in the dataset is
```{r}
sum(is.na(activitydata$steps))
```

Confirm that the other variables do not have NAs:
```{r}
sum(is.na(activitydata$date))
sum(is.na(activitydata$interval))
```

__Imputing strategy:__ use the interval average calculated in the above section and to replace NAs with interval average.

Calculate the total number of steps taken per day after imputing:
```{r results='asis'}
mergedact <- merge(activitydata, avgintervalsteps, by = "interval")
mergedact[is.na(mergedact$steps.x),]$steps.x <- mergedact[is.na(mergedact$steps.x),]$steps.y
mergedact <- mergedact[ order(mergedact[,3], mergedact[,1]), ]

dailystepsnona <- aggregate(formula = steps.x ~ date, data = mergedact, FUN = sum)
print(xtable(dailystepsnona), type = "html")
```

```{r}
hist(dailystepsnona$steps.x, breaks = 8)
```

Histogram of the total number of steps taken each day.

```{r}
imputeddailymean <- mean(dailystepsnona$steps.x)
imputeddailymedian <- median(dailystepsnona$steps.x)
```

The mean and median of the total number of steps taken per day are __`r imputeddailymean`__ and __`r imputeddailymedian`__ respectively.

After imputing values for NA, the mean and median have changed ever so slightly. However looking at the histogram, it looks like the variance in the data has decreased.

## Are there differences in activity patterns between weekdays and weekends?

```{r}

```
